---
output:
  pdf_document: default
  html_document: default
---
# Exploring dataset "impreseFVG" {-}
 

```{r include=FALSE}
# Local path, no need to display in knitted version
path = './../../_DATA/imprese'
library("tidyverse")
```
 
The dataset is organizes in a number of files; each file will be loaded in a different _data.frame_. 
 
```{r}
data.files <- list.files(path, pattern = ".csv$", recursive = TRUE)
print(paste("dataset contains",length(data.files), "files:"))
print(data.files)
```
## imprese
The core data identifying companies can be found in *t_imprese.csv* . 
```{r}
imprese <- read.csv( paste0(path,"/t_imprese.csv"), sep = "|")
str(imprese)
 
```
The attributes belong to different groups: 

- *metadata*:ï..fonte, mm_aaaa: 
- *identifier*: id_impresa,reg_imp_n, cf, piva, denominazione 
- *address*:prov,sede_ul,n.albo_art,reg_imp_sez  
- *type of company*: ng2
- *active status*: stato_impresa  
- *dates*:data_ini_at, data_cess_att, data_fall, data_liquid, data_cost, data_isc_ri, data_isc_rd,data_isc_aa,data_canc
- *employees*:  addetti_aaaa, addetti_indip, addetti_dip  
- *share capital*: capitale, capitale_valuta 
- *other attributes*: imp_startup, imp_femminile, imp_giovanile, imp_straniera,  imp_pmi, imp_sedi_ee, imp_eefvg 

```{r}
imprese$stato_impresa <- as.factor(imprese$stato_impresa)
df<-imprese %>% count(stato_impresa)  
ggplot(data=df, aes(x=stato_impresa, y=n)) +   geom_bar(stat="identity") + coord_flip()
imprese<- subset(imprese, stato_impresa =='ATTIVA')
nrow(imprese)
```
## Metadata
Metadata are generated by the pre-rpcessing algorithm and provide information about source and last update.
The two attributes (ï..fonte, mm_aaaa) are not relevant at this stage. 

## Identifiers
The following attributes are relevant: 
- denominazione: company name
- cf ("codice fiscale"): unique identifier, as factor (11 numbers or a string of 16 lettersa nd numbers)
- id_impresa: unique identifier, numeric.
Id and cf are unique, while company names are not and there are no missing values.

```{r}
imprese$cf <- as.factor(imprese$cf)
imprese$denominazione <- as.factor(imprese$denominazione)
# check missing calues
sum(is.na(imprese$denominazione)) + sum(is.na(imprese$cf)) == 0
# check duplicates in cf
length(unique(imprese$cf)) == length(imprese$cf)
# check duplicates in denominazione
uniqueNames <-length(unique(imprese$denominazione)) 
allNames<-length(imprese$denominazione)
print(paste("Company names are not a valid identifier for further analysis: the dataset contains", uniqueNames, "distinct names out of", allNames, "companies. There are", (allNames-uniqueNames), "duplicates."),quote = FALSE)

```
Other attributes (reg_imp_n,piva, n.albo_art,reg_imp_sez) are not relevant at this stage. 

## location
TODO
prov: province (GO, TS, UD, PN) >> factor FVG / ITA / EU
sede_ul: "SEDE" or "UL-n" >> factor SEDE = HeadOffice / UL = LocalUnit
LucalUnit = numeric 0 for HeadOffice, otherwise n
To be transformed in factors

- *type of company*: ng2
- *active status*: stato_impresa  
- *dates*:data_ini_at, data_cess_att, data_fall, data_liquid, data_cost, data_isc_ri, data_isc_rd,data_isc_aa,data_canc
- *employees*:  addetti_aaaa, addetti_indip, addetti_dip  
- *share capital*: capitale, capitale_valuta 
- *other attributes*: imp_startup, imp_femminile, imp_giovanile, imp_straniera,  imp_pmi, imp_sedi_ee, imp_eefvg 

```{r}
# company type: keep only the relevan ones for the scope of our research.
types <- read.csv( paste0(path,"/d_ng.csv"), sep = "|")
imprese$ng2           <- as.factor(imprese$ng2)
names(types)<-c("ngGroup", "ng2", "ngDescription")

df <- imprese  %>% count(ng2) 
df <- df %>% inner_join(types)
df <- df  %>% arrange(-n) %>% head(20)
df$ngDescription <- factor(df$ngDescription, levels = df$ngDescription) #lock factors to keep the same order in the bar plot
ggplot(data=df, aes(x=ngDescription, y=n)) +   geom_bar(stat="identity") + coord_flip()

```
Some company types are not relevant for our research, for example individual companies (DI) and other specified below. Dropping the corresponding dataframe rows drastically reduces the size of the data set
```{r}
notRelevant = c("DI", "AZ", "IR", "ER", "EP", "EN", "EM", "EL", "EE", "SM", "MA", "SZ", "LL", "AM", "AF")
toBeRemoved<-which(imprese$ng2 %in% notRelevant)
imprese2<-imprese[-toBeRemoved,]
print(nrow(imprese2))

df <- imprese2  %>% count(ng2) 
print(paste("The dataset contains ", nrow(df), "types of companies."), quote=FALSE)
df <- df %>% inner_join(types)
df <- df  %>% arrange(-n) 
df$ngDescription <- factor(df$ngDescription, levels = df$ngDescription) #lock factors to keep the same order in the bar plot
ggplot(data=head(df, 10), aes(x=ngDescription, y=n))  + geom_bar(stat="identity", color = "black", fill = "lightblue") + coord_flip() + ggtitle("Most common types") + geom_text(aes(x=ngDescription, y=n, label = n, hjust = -.2), color = "black") + ylim(NA, 22000)

```
 
<!-- ''' -->
<!--  ```{r} -->

<!-- #check if names areunique -->


<!-- imprese$prov          <- as.factor(imprese$prov) -->
<!-- imprese$imp_pmi       <- as.factor(imprese$imp_pmi) -->
<!-- imprese$imp_startup   <- as.factor(imprese$imp_startup) -->
<!-- imprese$imp_giovanile <- as.factor(imprese$imp_giovanile) -->
<!-- imprese$imp_femminile <- as.factor(imprese$imp_femminile) -->
<!-- imprese$imp_straniera <- as.factor(imprese$imp_straniera) -->
<!-- imprese$imp_sedi_ee   <- as.factor(imprese$imp_sedi_ee) -->

<!-- ``` -->



<!--  ```{r} -->
<!-- imprese$data_cost   <- as.Date(imprese$data_cost) -->
<!-- imprese$data_isc_aa <- as.Date(imprese$data_isc_aa) -->
<!-- imprese$data_isc_rd <- as.Date(imprese$data_isc_rd) -->
<!-- imprese$data_isc_ri <- as.Date(imprese$data_isc_ri) -->
<!-- imprese$data_ini_at <- as.Date(imprese$data_ini_at) -->
<!-- imprese$data_min    <- as.Date(imprese$data_isc_rd) -->

<!-- ``` -->


<!-- #imprese %>% mutate(data_min = coalesce(data_min, data_ini_at)) #verificare se funziona come previsto -->
<!-- #imprese %>% mutate(data_min = coalesce(data_min,data_isc_aa)) -->
<!-- #imprese %>% mutate(data_min = coalesce(data_min,data_isc_rd)) -->
<!-- #imprese %>% mutate(data_min = coalesce(data_min,data_isc_ri)) -->

<!--  ```{r} -->
<!-- imprese$data_min <- as.Date(imprese$data_min) -->
<!-- imprese <- imprese[!grepl('INATTIVA', imprese$stato_impresa),] -->

<!-- imprese <- imprese[complete.cases(imprese$data_min), ] #elimina le righe con NA in data_min -->

<!-- imprese$eta = as.numeric(as.Date("2022-01-01") - imprese$data_min) / 365  #calcola età -->
<!-- imprese <- imprese[complete.cases(imprese$eta), ] -->
<!-- eta <- imprese$eta -->


<!-- ``` -->



<!--  ```{r} -->
<!-- boxplot(imprese$eta, imprese$imp_giovanile) -->
<!-- ggplot(data=imprese, aes(x=eta, fill = imp_startup)) + geom_histogram(breaks=seq(0, 100, by=1)) -->
<!-- ``` -->



<!--  ```{r} -->
<!-- ggplot(data=imprese, aes(x=eta, fill = imp_femminile)) + geom_histogram(breaks=seq(0, 100, by=5)) -->

<!-- ``` -->


<!--  ```{r} -->
<!-- imprese2 <-subset(imprese,capitale_valuta == "EURO") -->
<!-- imprese2 <-subset(imprese2,capitale < 1e6) -->
<!-- imprese2 <-subset(imprese2,ng2 == "SP") -->


<!-- boxplot(imprese2$capitale) -->
<!-- qplot(imprese2$capitale, geom="histogram") -->
<!-- ``` -->


<!--  ```{r} -->


<!-- imprese<- subset(imprese2, select = c('eta', 'cf', 'id_impresa') ) -->
<!-- ids <- unique(imprese2$id_impresa) -->
<!-- localiz <- read.csv("t_localizz.csv", sep = "|") -->
<!-- localiz <- localiz[localiz$id_impresa %in% ids,] -->
<!-- localiz <- subset(localiz, select = c('id_impresa', 'id_localiz')) -->
<!-- locs <- unique(localiz$id_localiz) -->
<!-- codici <- read.csv("t_codici.csv", sep = "|") -->
<!-- codici <- codici[codici$id_localiz %in% locs,] -->
<!-- codici <- subset(codici, select = c('id_localiz', 'ateco')) -->

<!-- codici = merge(codici, localiz, by = "id_localiz") -->

<!-- imp <- data.frame("company"=ids) -->
<!-- imp$eta <- imprese$eta -->
<!-- imp$ateco <- NA -->

<!-- #aggiunge al data.frame una lista di codici ateco -->
<!-- for (impresa in imp$company){ -->
<!--         temp = as.list(subset(codici, id_impresa == impresa)) -->
<!--         unique_nace_codes <- unique(temp$ateco) -->
<!--         rownum = which(imp$company == impresa) -->
<!--         imp$ateco[rownum] <- list(unique_nace_codes) -->
<!-- } -->

<!-- " -->
<!-- '''" -->

