---
output:
  pdf_document: default
  html_document: default
---
# Exploring dataset "bilanciFVG"
This section is dedicated to load and preprocess financial statement data for the dataset *imprese-fvg*. 
The relevant file is "_DATA/imprese-fvg/bilanci-fvg.csv".

```{r include=FALSE}
# Local path, no need to display in knitted version
pathRawData = './../../_data/raw/'
pathTidyData = './../../_data/tidy/'
library("tidyverse")
```
 
The relevant file is *bilanci-fvg.csv*. Each observation is a summary of balance sheet data (bsd) of a company (identified by *cf*) for a given year. Column labels need some improvement to remove whitespaces and possibly short english names.
 

```{r}
bsd <- read_delim( paste0(pathRawData,"imprese/bilanci-fvg.csv") ) 
spec(bsd) # tydiverse for str(companies)

```

```{r}
bsd <- bsd %>% 
        rename(year = anno) %>%
        rename(totEquity = `Totale patrimonio netto`) %>%
        rename(totAssets = `Totale attivo`) %>%
        rename(totIntang = `Totale Immobilizzazioni immateriali`) %>%
        rename(staffCost = `Totale Costi del Personale`) %>%
        rename(turnover  = `Ricavi delle vendite`) %>%
        select(cf,year,turnover, totAssets, totIntang, staffCost  )  

```


```{r}
bsd <- bsd %>% 
  mutate(across(everything(), gsub, pattern = "[.]", replacement = "")) %>%
  mutate(across(everything(), gsub, pattern = ",", replacement = ".")) %>%
  mutate(across(.cols = 2:6, .fns = as.numeric))

```



```{r}
bsd %>% write_csv(paste0(pathTidyData,"bsd.csv"))
```


 
# ```
There are 18 columns but in this project we will use only 4, namely "cf", "year", revenues" and "staff cost". Data should be loaded as string and then converted taking into account some issues with format of numerical variables. 
 
To convert bsd$revenues and bsd$staffcost to numbers, we need to remove the "." used as thousand separators, and replace "," with "." as a decimal separator.

We will focus the analysis on a list of companies that are tenats at Area Science Park. The list is available in the file "data/imprese-fvg/area-tenants.csv" so we can load it in al list ("filter") and use it to subset *bsd*.

```{r}

tenants <- read_delim( paste0(pathRawData,"area-science-park/tenants.txt") ) %>%
  select(cf)  

tens = c(tenants$cf)
bsd_tenants <- bsd %>% subset(cf %in% tenants$cf) %>% 
                mutate(cf = as.factor(cf)) %>% drop_na() 
  
```
The variable bsd$revenues spans from 0 to 1e9, so it is more convenient to work with log10

```{r}
library(ggplot2)
library(ggpubr)
bsd3 <- bsd %>% subset(turnover > 1000) %>% subset(year = 2019)
bsd3$logturnover <- log10(bsd3$turnover)
# hist(bsd$turnover)
# hist(bsd$logturnover)
h1 <- ggplot(bsd3, aes(x=turnover)) + geom_histogram(color="black", fill="red")
h2 <- ggplot(bsd3, aes(x=logturnover)) + geom_histogram(color="black", fill="green", aes(y=..density..))+  geom_density(alpha=.2, fill="#FF6666") 
figure <- ggarrange(h1, h2,labels = c("linear", "log"), ncol = 2, nrow = 1)
figure

```
 
 
 
```{r}
bsd_tenants$logturnover <- log10(bsd_tenants$turnover)

tmp <- bsd_tenants  %>%
              subset(year >= 2016) %>% 
              mutate(year = as.factor(year))

figure <- ggplot(tmp, aes(x=logturnover,fill=year)) + geom_density(alpha=.2)  
figure

```
