---
output:
  pdf_document: default
  html_document: default
---
# Exploring dataset "CO-FVG" 


```{r include=FALSE}
# Local path, no need to display in knitted version

pathRawData = 'C:/Users/morea/OneDrive - Area Science Park/General - iifvg lavoro/DatiElaborati/Riservati/'
pathTidyData = './../../_data/tidy/'
library("tidyverse")
data.files <- list.files(pathRawData, pattern = "dati_*", recursive = TRUE )
```


The original data is organized in `r length(data.files)` files: `r data.files`. 
> TODO Currently, data exploration phase is focused on only one of the files above. Should extend it to all files using a for loop and appending results to a data.frame.

```{r}
empl <- read_delim( paste0(pathRawData,"dati_2018.csv")) 

features <-names(empl)
some_features <- c("CF","anno","eta","genere","iso3","professione","qualifica","saldo")


```

 There are `r length(features)` features available: `r features`. For the purpose of data exploration we will focus only on the following: `r some_features`.


```{r}

empl <- empl %>%  
  select( one_of(some_features) ) %>%
  rename( year = anno)
 
empl_flows <- empl %>% select( c(CF, saldo, year)) %>%  
  mutate(hf = factor(saldo))%>% 
  mutate(hf=recode(hf,`-1`="fired",`1`="hired"))%>%
  group_by(CF,hf, year) %>% 
  summarize(hiredfired=  sum(saldo) ) %>% 
  pivot_wider( names_from = hf, values_from = hiredfired) %>%
  replace(is.na(.), 0) %>%
  mutate(turnover = hired-fired) %>%
  mutate(net = hired+fired) 
 
```


```{r}
employees_hired = log10(empl_flows$hired)
employees_fired=  log10(-empl_flows$fired)
ggplot(empl_flows, aes(x=employees_hired, y=employees_fired))+
    geom_point()



```



```{r}
employees_hired = (empl_flows$hired)
employees_fired=  (-empl_flows$fired)
ggplot(empl_flows, aes(x=employees_hired, y=employees_fired))+
    geom_point()



```
 > TODO import, calculate net saldo and turnover, divide companies in quartiles

> TODO improve formatting tables with library(kableExtra) %>% kable()


```{r}
empl_flows %>% write_csv(paste0(pathTidyData,"empl_flows.csv"),)
```

