---
title: "R tidyverse"
output: pdf_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE)  # avoid warning messages in PDF
path = './../../_DATA/imprese' #local path to datasets
```


# Testind tidyverse

reference website: https://www.tidyverse.org/index.html 

reference e-book: https://r4ds.had.co.nz/

Tidyverse is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures.

Install the complete tidyverse with:
```{r}
#install.packages("tidyverse")
library(tidyverse)
```

Basic tidyverse packages include dyplr, tibble and ggplot2.
Additional tidyverse packages that help with general programming challenges: purr (loops), and magrittr (pipe operator)

## magrittr %>% pipe operator %>%
magrittr provides the pipe operator %>% used throughout the tidyverse. It also provide a number of more specialised piping operators (like %$% and %<>%) that can be useful in other places.
basic meaning of piping: 

* x %>% f is equivalent to f(x)
* x %>% f(y) is equivalent to f(x, y)
* x %>% f %>% g %>% h is equivalent to h(g(f(x)))

In the following example we calculate a new variable (addetti_tot), filter only the observations that have more than 500 addetti_tot and province in FVG, finally sort.

```{r}
path = './../../_DATA/imprese'
#load a dataset
large_companies <- read_delim( paste0(path,"/t_imprese.csv")) %>%
                   transform(addetti_tot = addetti_indip + addetti_dip  ) %>%
                   subset(addetti_tot > 500) %>%
                   subset(prov %in% c("GO","TS","UD","PN")) %>%
                   arrange(-addetti_tot) %>%
                   select(c("denominazione", "addetti_tot")) 
large_companies %>% head(5)
```

```{r}
# define classes and count the number of companies by class
classes <- large_companies %>%
            mutate(groups = case_when(    addetti_tot >= 2000 ~ 'XL',
                   addetti_tot >= 1000  & addetti_tot <= 1999 ~ 'L',
                   addetti_tot >= 500   & addetti_tot <= 999  ~ 'M')) %>%
           group_by(groups) %>%
           summarise(n())
classes  

```


```{r}
tmp <-read_delim( paste0(path,"/t_imprese.csv"))
```

